Airspace configuration decision based on Monte-Carlo tree search

### 2018-04-18 article/article.tex
(+) Continued

The bandit problem is applied to MCTS via the Upper Confidence Tree algorithm.
It uses the UCB for Upper Confidence Bound as tree policy (introduced
in~\cite{kocsis2006bandit}). Let \(v\) be the node from which an action must be
selected to go deeper in the tree, \(a\) an action available from \(v\),
\(\mu_{v,a}\) the expected outcome from the node obtained by applying action
\(a\) on \(v\), \(T_v\) the number of simulations carried out from node \(v\)
(which includes any simulation from nodes in any subtree of \(v\)) and
\(T_{v,a}\) the number of simulations carried out from the node resulting from
the action \(a\) applied to node \(v\), \(\beta\) a constant. The selected node
is thus the one maximising
\begin{equation}
  \text{UCB} = \mu_{v,a} + 2 \beta \sqrt{\frac{2 \log T_v}{T_{v,a}}}
\end{equation}

\paragraph{Expansion}
The above step is carried out only if, considering a node \(F\), a simulation
has been run through each child available from \(F\). In the case where one or
more children haven't been explored yet, one of the unexplored children is
chosen and a simulation is started from it.
\begin{algorithm}
  \caption{UCT algorithm}\label{alg:tree_pol}
  \begin{algorithmic}
    \Function{TreePolicy}{path of nodes $\pi$}
    \State{} $v \gets$ head of $\pi$
    \State{} \Call{Deploy}{$v$}
    \If{$v$ is terminal}
    \State{}\Return{$\pi$}
    \Else{}
    \If{$v$ is not fully expanded}
    \State{}\Return{\Call{Expand}{$v$} $\cons \pi$}
    \Else{}
    \State $f \gets$ \Call{BestChild}{$v$}
    \State{}\Return{%
      \Call{TreePolicy}{$f \cons \pi$}
    }
    \EndIf{}
    \EndIf{}
    \EndFunction{}
  \end{algorithmic}
  \begin{algorithmic}
    \Function{BestChild}{$v$}
    \State{} \Return{$\argmax\{\Call{UCB}{v'} |
    v' \text{\,children of\,} v\}$}
    \EndFunction{}
  \end{algorithmic}
\end{algorithm}
\begin{algorithm}
  \caption{%
    Expansion (\(p\colon \mathcal{S} \times \mathcal{A} \to \mathcal{S}\)
    is the transition function)
  }
  \begin{algorithmic}
    \Function{Expand}{node $v$}
    \State{} $\ell \gets \{p(v, a) | a\in\mathcal{A}(v), T_{v,a} = 0\}$
    \State{}\Return{random element from $\ell$}
    \EndFunction{}
  \end{algorithmic}
\end{algorithm}

\section{Model}

>>>>

The bandit problem is applied to MCTS via the Upper Confidence Tree algorithm.
It uses the UCB for Upper Confidence Bound as tree policy (introduced
in~\cite{kocsis2006bandit}). Let \(v\) be the node from which an action must be
selected to go deeper in the tree, \(a\) an action available from \(v\),
\(\mu_{v,a}\) the expected outcome from the node obtained by applying action
\(a\) on \(v\), \(T_v\) the number of simulations carried out from node \(v\)
(which includes any simulation from nodes in any subtree of \(v\)) and
\(T_{v,a}\) the number of simulations carried out from the node resulting from
the action \(a\) applied to node \(v\), \(\beta\) a constant. The selected node
is thus the one maximising
\begin{equation}
  \text{UCB1} = \mu_{v,a} + 2 \beta \sqrt{\frac{2 \log T_v}{T_{v,a}}}
\end{equation}

While the previous equation is well suited for two players games, it can be
tweaked to improve its efficience in one player games or sequencing problems. An
alternative using the standard deviation is proposed in~\cite{sebag2010fuse}
called the \emph{UCB1-tuned} equation
\begin{equation}
  \mu_{v,a} +
  \sqrt{%
    \frac{\beta\log T_v}{t_{v,a}}
    \min\left( \frac{1}{4}, \sigma^2_{v,a} +
    \sqrt{%
      \frac{2 \log T_v}{T_{v,a}}
    }\right)
  }
\end{equation}
\begin{algorithm}
  \caption{UCT algorithm}\label{alg:tree_pol}
  \begin{algorithmic}
    \Function{TreePolicy}{path of nodes $\pi$}
    \State{} $v \gets$ head of $\pi$
    \State{} \Call{Deploy}{$v$}
    \If{$v$ is terminal}
    \State{}\Return{$\pi$}
    \Else{}
    \If{$v$ is not fully expanded}
    \State{}\Return{\Call{Expand}{$v$} $\cons \pi$}
    \Else{}
    \State $f \gets$ \Call{BestChild}{$v$}
    \State{}\Return{%
      \Call{TreePolicy}{$f \cons \pi$}
    }
    \EndIf{}
    \EndIf{}
    \EndFunction{}
  \end{algorithmic}
  \begin{algorithmic}
    \Function{BestChild}{$v$}
    \State{} \Return{$\argmax\{\Call{UCB1-tuned}{v'} |
    v' \text{\,children of\,} v\}$}
    \EndFunction{}
  \end{algorithmic}
\end{algorithm}

\paragraph{Expansion~\ref{alg:expansion}}
The above step is carried out only if, considering a node \(v\), a simulation
has been run through each child available from \(v\). In the case where one or
more children haven't been explored yet, one of the unexplored children is
chosen and a simulation is started from it.
\begin{algorithm}
  \caption{%
    Expansion (\(p\colon \mathcal{S} \times \mathcal{A} \to \mathcal{S}\)
    is the transition function)
  }\label{alg:expansion}
  \begin{algorithmic}
    \Function{Expand}{node $v$}
    \State{} $\ell \gets \{p(v, a) | a\in\mathcal{A}(v), T_{v,a} = 0\}$
    \State{}\Return{random element from $\ell$}
    \EndFunction{}
  \end{algorithmic}
\end{algorithm}

\subsubsection{Simulation and backpropagation}
\paragraph{Simulation}
Once a node has been expanded, random nodes are chosen successively until a
terminal state is found. The cost of the resulting path, from the root to the
node is then evaluated and backpropagated to all the ancestors of the expanded
node.

\subparagraph{Heuristic} One might want to bias the randomness while choosing
nodes. This would imply using a heuristic which has to be able to discriminate
a node between its siblings.

\paragraph{Backpropagation}
The backpropagation consists in updating the values required to carry out the
tree policy. The values must be updated incrementally since the backpropagation
function has the current value of the parameters and the result of the
simulation as parameters. Thus, for UCB1 equation, the expected reward (mean)
and the total count are updated this way
\begin{algorithm}
  \caption{UCB1 backpropagation}
  \begin{algorithmic}
    \Procedure{Backpropagate}{path $\pi$, $r$}
    \State{} \(\delta \gets\ r - \mu_v\)
    \State{} \(\mu_v \gets \mu_v + \frac{\delta}{T_v + 1}\)
    \State{} \(T_v \gets T_v + 1\)
    \EndProcedure{}
  \end{algorithmic}
\end{algorithm}

For the UCB1-tuned, the standard deviation has to be computet. It results in
algorithm~\ref{alg:backpropagate-tuned}
\begin{algorithm}
  \caption{UCB1-tuned backpropagation}\label{alg:backpropagate-tuned}
\end{algorithm}
\section{Model}
